---
title: "ESM 244 | Lab 6 | Spatial"
author: "Hanna Buechi"
date: "2/14/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Goals: 

- Read in spatial data (shapefiles)
- Simplify polygons (st_simplify)
- Add and transform projection (st_transform and st_crs)
- Create several finalized-ish maps in R
- Join spatial data (st_join)
- Find spatial intersections (st_intersect)
- Interactive plots with tmap
- Updating color schemes, base layers, etc.
- Plotting spatial lines
- Creating sf data from lat/lon coordinates

Awesome resource: 
*Geocomputation in R* by Robin Lovelace, available online: 
<https://geocompr.robinlovelace.net/>

```{r}
library(tidyverse)
library(sf)
library(tmap)
library(leaflet)
library(ggrepel)
library(ggspatial)
library(RColorBrewer)
library(raster)
```

Important files in a spatial dataset:

.shp - geometries
.dbf - attributes
.prj - projections

Not all datasets will come with all of these. For example, we may have to create a projection file.


###Example 1. Dams in California

This data includes point location of dams and attributes associated with them: ownership, capacity, height of dam, downstream hazard, etc.

####Read in data
```{r eco_regions}

ca_eco <- read_sf(".", layer = "ca_eco") %>% # "." means get this from the current working directory
  dplyr::select(US_L3NAME) %>% # if we view ca_eco, we see those column titles - this is just one of the columns for all of the attributes; sticky geometries!!!
  rename(Region = US_L3NAME) %>% 
  st_simplify(dTolerance = 100) %>% # larger tolerance --> lower spatial resolution to collapse more points onto single line; 100 m; makes running code faster, but should only be done at an appropriate scale
  st_transform(crs = 4326) # transform existing projection system (st_crs() if there was no coordinate system); 4326 = WGS?

st_crs(ca_eco) # check projection system

```

```{r county}

ca_counties <- read_sf(".", layer = "california_county_shape_file")

st_crs(ca_counties) = 4326 # so easy to transform projection system!!

```

```{r dams}

ca_dams <- read_sf(".", "California_Jurisdictional_Dams") %>% 
  rename(Condition = Condition_) # change column name

ca_dams$Condition <- fct_relevel(ca_dams$Condition, "Fair", "Satisfactory", "Unsatisfactory", "Poor") # want an order that isn't alphabetical; important for data visualization!

```


MAPPPPSSSSSSSSSSSS
```{r base_plot}

plot(ca_eco) # kept only one attribute so only 1 map
plot(ca_counties) # there are many attributes because we kept them all

```


Make a map with ggplot!
```{r ggplot}

# set number of colors for R Color Brewer

color_count <- 13 # need to know how many colors we need...

my_colors <- colorRampPalette(brewer.pal(10, "Set2"))(color_count) # use 10 colors from Set2 with number of color_count colors # using colorRampPalette overrides the error

ggplot(ca_eco) +
  geom_sf(aes(fill = Region),
          color = "NA", # borders around polygons
          show.legend = FALSE) + # gets rid of legend
  scale_fill_manual(values = my_colors) +
  ggtitle("Dams in California") +
  geom_sf(data = ca_counties, # add the county borders
          fill = NA,
          color = "gray 30",
          size = 0.1) +
  geom_point(data = ca_dams,
             aes(x = Longitude, y = Latitude),
             size = 0.5,
             color = "gray10",
             alpha = 0.5) + # can color dams by attributes, too
  coord_sf(datum = NA) +
  theme_light()
  
  

```


###Example 2. Dams in the Sierra Nevada eco-region

```{r subset_sn}

sn <- ca_eco %>% 
  filter(Region == "Sierra Nevada") %>% 
  st_join(ca_dams) # join wherever data from dams and eco regions joins up

ggplot(sn) + # base ggplot data, but data won't "show up" until geom_sf()
  geom_sf(data = ca_counties, fill = "wheat3", color = "NA") + # map of California with county polygons
  geom_sf(fill = "lemonchiffon4", color = "NA") + # don't need to remind it sn # polygons on polygons
  geom_point(aes(x = Longitude, y = Latitude), # don't need to remind it sn + points on polygons
             size = 0.5,
             color = "red4")

```


###Example 3. Eco-regions for Santa Barbara County
```{r}

sb <- ca_counties %>% 
  filter(NAME == "Santa Barbara") # each row is a different polygon and Santa Barbara Co has islands!

eco_clip <- st_intersection(ca_eco, sb) # clip ca_eco by the bounds of sb

ggplot(eco_clip) +
  geom_sf(data = ca_counties, # still using basemap of california counties
          fill = "grey90", 
          color = "grey80",
          size = 0.2) +
  geom_sf(aes(fill = Region), 
          color = "NA") +
  scale_fill_manual(values = c("darkolivegreen2","darkolivegreen","gold2")) +
  coord_sf(xlim = c(-121, -119), ylim = c(33.5, 35.5)) + # limit the coordinates (allison did trial-and-error)
  geom_point(aes(x = -119.6982, y = 34.4208), size = 2) + # Santa Barbara
  geom_text(x = -119.6982, y = 34.35, label = "Santa Barbara") +
  theme(legend.position = c(0.5,0.15)) + # Move the legend, location in a non-cropped version of the graph... mess around with it...
  labs(x = "", y = "", title = "Santa Barbara County Eco-Regions") +
  theme_light()

```

###Example 4. Intro to interactive plots with tmap

```{r}

# make a tmap object that we will plot interactively
tmap_mode("view") # run this first so that interactive map comes up
map_sb_eco <- tm_shape(eco_clip) + # take this information and make it into an interative tmap object
  tm_fill("Region", palette = "RdPu", alpha = 0.5) + # fill the polygons
  tm_shape(ca_counties) +
  tm_borders() # need this with tm_shape
  
# Example with a different basemap
# leaflet::providers to see list of basemaps

test_basemap <- tm_basemap("CartoDB.DarkMatter") +
  tm_shape(eco_clip) +
  tm_borders(col = "white") # we didn't put counties in

```

###Example 5. Fault lines

```{r}

fault_lines <- read_sf(".", layer = "GMC_str_arc") %>% 
  st_transform(crs = 4326) %>% 
  separate(LTYPE, into = c("syn_ant", "certainty", "direction"), sep = ",") # I've done this before!

plot(fault_lines)

ggplot() + # faults!!!!!
  geom_sf(data = ca_counties, fill = "black", color = "NA") +
  geom_sf(data = fault_lines, aes(color = syn_ant)) +
  theme_light()

# fault lines in SB County:

sb_faults <- fault_lines %>% 
  st_intersection(sb) # find fault lines in sb

ggplot() +
  geom_sf(data = sb) +
  geom_sf(data = sb_faults, aes(color = syn_ant)) # those fault lines in SB county

tm_basemap("CartoDB.DarkMatter") +
  tm_shape(sb) +
  tm_borders(col = "gray50", lwd = 2) +
  tm_shape(sb_faults) +
  tm_lines(col = "syn_ant", palette = c("orange", "purple"), lwd = 2)

```


###Example 6. Faceted maps

```{r}

ggplot() +
  geom_sf(data = ca_counties, fill = "black", color = "NA") +
  geom_sf(data = fault_lines, aes(color = syn_ant)) +
  facet_wrap(~syn_ant) # type of faults

# more examples in lab key

```


###Example 7. Making spatial points

```{r sensitive_sites}

ca_sites <- read_csv("cadfw_sensitive_sites.csv")
sites_sf <- st_as_sf(ca_sites, coords = c("LONDD", "LATDD"), crs = 4326) # telling R that this is spatial data

# now this is spatial! Makes sticky geometries!

ggplot() +
  geom_sf(data = ca_counties, fill = "gray40") +
  geom_sf(data = sites_sf, aes(color = PRI_CODE), size = 0.3) # notice how the labels are in alphabetical order

```


###Example 8. Chloropleth of CA counties by NUMBER of dams in each

```{r}

intersection <- st_intersection(x = ca_dams, y = ca_counties)

dams_per_county <- intersection %>% 
  group_by(NAME) %>% # count all of those dams in the county
  tally()

ca_tot <- ca_counties %>% 
  st_join(dams_per_county) %>% 
  dplyr::select(NAME.x, n) %>%
  rename(name = NAME.x)

# Reassign NA values to zero:
ca_tot$n[is.na(ca_tot$n)] <- 0 # this is important because there were some counties that didn't have dams or some dams that didn't have counties

ggplot() +
  geom_sf(data =ca_tot, aes(fill = n), size = 0.2) +
  theme_minimal() +
  scale_fill_continuous(low = "yellow", high = "red")





```

































